---
title: "rlernen"
author: "Johannes Titz"
date: "2025-04-02"
output: github_document
---

## Einführung

Dies ist das Repositorium für den R-Kurs *rlernen*. Es gibt zwei Wege den Kurs zu bearbeiten: 

(1) online über https://rlernen.de (benötigt keine Installation von R, es kann sofort losgehen)
(2) lokal mit einer eigenen R-Installation, geht auch offline und wird häufiger aktualisiert

Wenn Du Weg 2 wählst, gehe folgende Schritte:

- installiere R
- installiere RStudio (optional, macht die Sache aber etwas einfacher)
- installiere das Paket rlernen

Für die Installation von R und RStudio folge diesem Tutorial: https://jjallaire.shinyapps.io/learnr-tutorial-00-setup/

Nun kannst Du *rlernen* installieren, wofür Du allerdings vorab remotes brauchst:

```{r, message=F}
install.packages("remotes")
remotes::install_github("johannes-titz/rlernen")
```

Updates erfolgen über:

```{r, message=F}
remotes::install_github("johannes-titz/rlernen")
```

Um die Tutorials zu starten kannst Du in RStudio rechts in der Ecke auf "Tutorial" klicken und runterscrollen bis Du die rlernen-Tutorials siehst. Alternativ kannst Du in der Konsole einfach `learnr::run_tutorial("tag1", "rlernen")` ausführen, was auch direkt aus R heraus geht. Ändere "tag1" zum Tutorial, was Du bearbeiten möchtest.

## Hausaufgaben

Wenn Du diesen Kurs im Rahmen des Studiums an der TU Chemnitz machst, kannst Du Hausaufgaben zu jedem Thema abgeben und Feedback erhalten. Um die Datei zur jeweiligen Sitzung zu öffnen, führe folgenden Code aus:

```{r eval=FALSE}
library(rlernen) # laden des Pakets
homework(1) # 1: erste Sitzung, ändere die Zahl entsprechend
```

Die Abgabe der Hausaufgaben erfolgt über OPAL:

https://bildungsportal.sachsen.de/opal/auth/RepositoryEntry/8420720649/CourseNode/91305593842947

<details><summary><b>Notes for self-hosting this tutorial</b>
</summary>
This repository contains the learnr-based tutorial *rlernen*. It is primarily intended for personal use but may also be useful for others who wish to modify or host the *rlernen* tutorial.

## Important Note

There are unresolved caching issues with *learnr*. Caching is disabled for `tag2` and `tag5` due to the following problems:
- `tag2` fails to render entirely, causing excessive loading times.
- `tag5` disrupts navigation.

## Version 2.0 (September 2023)

- Updated all tutorials.
- Added a new section on reading and writing data.
- Created a custom help function.
- Reduced dependencies, decreasing image size from 3.62 GB to 1.89 GB.

## To-Do List

- Optimize and test the Dockerfile (use `glue_sys_reqs` from *mimosa*).
- Develop an R/RStudio installation tutorial similar to Posit's, and link it at the end of *Day 1* (suggested by Alexandra).

## Procedure

### Docker

After modifying `.Rmd` files, push changes to GitHub, pull them on the server, and rebuild the Docker image:

```{sh eval=FALSE}
sudo docker build . -t rlernen
```

Test the container (modify the file to see all tutorials):

```{sh eval=F}
docker run --rm -p 4000:3838 rlernen R -e "rmarkdown::run('tag2.Rmd')"
```

- **Why is port 4000 used?** ShinyProxy expects this port for proper communication.
- **`--rm` flag:** Ensures the container is removed after execution to prevent clutter in the Docker container list.
- **`R -e` flag:** Executes an R command directly from the command line.

## Website Deployment

**Project:** `programming/rkurs_website`

1. Modify `index.html` as needed.
2. Run the script `create_nav.R` to update navigation.
3. Upload the updated site to the server using:

```{sh eval=F}
rsync -arvz --stats --progress --itemize-changes -e 'ssh -p 22' \
  _site/ johannes@method.hsw.tu-chemnitz.de:/var/www/rlernen_v2.0.de
``` 

</details>
